---
title: "Practical Machine Learning"
author: "Me"
date: "4 7 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Downloading and cleaning data

Data can be downloaded from provided URLs

```{r}

# Load necessary libraries

library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)

#Download dataset

training <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
testing <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

#Create a partition on the training set

part <- createDataPartition(training$classe, p = 0.7, list = FALSE)
TrainSet <- training[part, ]
TestSet <- training[-part, ]

```

We will be using 70% of the data for training purposes and the remaining 30% for testing purposes.

```{r}
dim(TrainSet)
dim(TestSet)

```


Out of the 160 variables, we can exclude those wo contains NA or that are approx. zero and also the 5 used for ID.

```{r}
# remove NA
TrainSet <- TrainSet[ , colSums(is.na(TrainSet)) == 0]
TestSet <- TestSet[ , colSums(is.na(TestSet)) == 0]

# remove variables with approx zero variance
TrainSet<- TrainSet[, -nearZeroVar(TrainSet)]
TestSet <- TestSet[, -nearZeroVar(TestSet)]

# remove the first 5 columns -ID only

TrainSet<- TrainSet[ , -c(1:5)] 
TestSet <- TestSet[ , -c(1:5)]

```

We have now reduced our variables from 160 to 54

```{r}
dim(TrainSet)
dim(TestSet)
```

## Training Model

Let´s now try to model the cleaned data; we now from the course that random forest usually perform very well in this kind of inquiries so we will fit it to our datasets. The confusion matrix provide us a better overview about accuracy of the different models.


```{r}
set.seed(12345)
ctr <- trainControl(method = "cv", number = 3, verboseIter = FALSE)
RF <- train(classe ~., data = TrainSet, method = "rf", trControl = ctr)
RF$finalModel
```

```{r}
#prediction on test dataset
predRF <- predict(RF, TestSet)
#confusion matrix
confMatRF <- confusionMatrix(predRF, TestSet$classe)
confMatRF
```

With 99,75% accuracy we can say that our model is really good.

## Apply the model to the test data

We can finally apply the random forest model to the 20 cases provided as test-dataset-

```{r}
TEST <- predict(RF, testing)
TEST

```



